Towards Generic Parallel Programming in Computer Science Education with Kokkos

### 2020-10-07 src/6.Backend.tex
EduHPC @ SC2020 Camera Ready

A simplified version of a partially specialized class for a parallel for-loop implementation in the OpenMP backend is shown in Figure\ref{fig:KokkosExampleOMPBackEnd}. The ~\emph{parallel\_for} function instantiates a partial specialization of the ~\emph{ParallelFor} class and calls its ~\emph{execute} function. That function is then responsible to implement the Kokkos parallel pattern. In this backend, the parallel implementation is generated by using OpenMP pragma annotations and a by invoking a compatible compiler when the library is compiled. Figure~\ref{fig:stack} gives an architectural overview of Kokkos and lists important abstractions as well as the currently supported backends.
\begin{figure}
\begin{small}
\begin{minted}[fontsize=\small]{c++}
template <class FunctorType>
class ParallelFor<FunctorType,
              RangePolicy<OpenMP>> {
  const FunctorType functor;
  const RangePolicy<OpenMP> policy; 
public:
  void execute() const {
    #pragma omp parallel for
    for(int i=policy.begin();
            i<policy.end();
            ++i)
      functor(i);
  }
};
template <class FunctorType,class PolicyType>
void parallel_for(std::string label, 
    const PolicyType& policy,
    const FunctorType& functor) {
  // Call profiling tool hooks with label
  ParallelFor<FunctorType,PolicyType> 
    pf{functor,policy};
  pf.execute();
}
\end{minted}
\end{small}
\caption{This code shows the backend implementation of a parallel loop in the Kokkos programming model. It shows a particular implementation of a generic pattern as a templated class to the OpenMP execution space. Also, it shows the use of OpenMP pragma annotations that put platform compilers in charge to generate parallel code for the underlying hardware architecture.}
\label{fig:KokkosExampleOMPBackEnd}
\end{figure}

>>>>

A simplified version of a partially specialized class for a parallel for-loop implementation in the OpenMP backend is shown in Figure~\ref{fig:KokkosExampleOMPBackEnd}. The ~\emph{parallel\_for} function instantiates a partial specialization of the ~\emph{ParallelFor} class and calls its ~\emph{execute} function. That function is then responsible to implement the Kokkos parallel pattern. In this backend, the parallel implementation is generated by using OpenMP pragma annotations and a by invoking a compatible compiler when the library is compiled. Figure ~\ref{fig:stack} gives an architectural overview of Kokkos and lists important abstractions as well as the currently supported backends.
\begin{figure}
\begin{small}
\begin{minted}[fontsize=\small]{c++}
template <class FunctorType>
class ParallelFor<FunctorType,
              RangePolicy<OpenMP>> {
  const FunctorType functor;
  const RangePolicy<OpenMP> policy; 
public:
  void execute() const {
    #pragma omp parallel for
    for(int i=policy.begin();
            i<policy.end();
            ++i)
      functor(i);
  }
};
template <class FunctorType,class PolicyType>
void parallel_for(std::string label, 
    const PolicyType& policy,
    const FunctorType& functor) {
  // Call profiling tool hooks with label
  ParallelFor<FunctorType,PolicyType> 
    pf{functor,policy};
  pf.execute();
}
\end{minted}
\end{small}
\caption{This code shows the backend implementation of a parallel loop in the Kokkos programming model. It shows a particular implementation of a generic pattern as a templated class to the OpenMP execution space. Also, it shows the use of OpenMP pragma annotations that put platform compilers in charge to generate parallel code for the underlying hardware architecture.}
\label{fig:KokkosExampleOMPBackEnd}
\end{figure}

### 2020-10-07 src/9.Teaching.tex
EduHPC @ SC2020 Camera Ready

Our presentation and tutorial material is divided into 8 units that gradually roll out all features and use-cases of the Kokkos programming model. Our tutorial material covers approximately 8 two-hour lectures, where each lecture summarizes the key learning points of the previous lecture at the beginning, summarizes the current lecture at the end, and gives a preview on the upcoming lecture. This creates continuity, improves the attendee's orientation in the subject, and spurs interest. Each lecture is concluded with a call for experimentation. Each lecture has a prepared hands-on exercise that can be used either in-lined during presential lectures or given out as homework for individual experimentation. During online lectures, the presenter explains the solution of the homework through screen-sharing and repeats important concepts. We would like to point out that each homework assignment contains the solution. Providing a solution lowers the barrier of experimentation and helps impatient students.

Online lectures tend to restrict possible interaction between the presenter and attendees to chats. In turns out useful to assign assistant presenters to actively participate in the chat during the presentation. This helps to resolve questions and support audience engagement. Since some chat conversations tend to fall short, it is a good practice to save the chat transcript and to publish the transcript together with refined responses on-line where they can be accessed by attendees for reference. To help interested attendees, we offer consulting hours. Consulting hours are hosted as online telephone conferences where attendees can call in and ask further questions regarding lecture content or homework.

@@

A recent online course\cite{KOKKOS_LECTURE} has been publicly announced through social media and counted approximately 200 attendees. The majority of attendees are familiar with at least one parallel programming model and have a background in scientific computing. The listing in Figure~\ref{fig:lectureOutline} shows the structure of our online lecture.

Question asked by attendees during the recent online course showed that they did understand the relevance as well as key concepts. Questions were mostly related to details in behavior, differences to other programming models and best practices. We would like to direct the interested reader to the our online website where the Q\&A sessions are listed in detail.

>>>>

Our presentation and tutorial material is divided into 8 units that gradually roll out all features and use-cases of the Kokkos programming model. Our tutorial material covers approximately 8 two-hour lectures, where each lecture summarizes the key learning points of the previous lecture at the beginning, summarizes the current lecture at the end, and gives a preview on the upcoming one. This creates continuity, improves the attendee's orientation in the subject, and spurs interest. Each lecture is concluded with a call for experimentation.

Each unit has a prepared hands-on exercise that can be used either in-lined during presential classes or given out as homework for individual experimentation. During the online presentation, the presenter explains the solution of the homework through screen-sharing and repeats important concepts. We would like to point out that each homework assignment contains the solution in a separate directory. Providing a solution lowers the barrier of experimentation and helps impatient students.

Online classes tend to restrict possible interaction between the presenter and attendees to chats. In turns out useful to assign assistant presenters to actively participate in the chat during the presentation. This helps to resolve questions and support audience engagement. Since some chat conversations tend to fall short, it is a good practice to save the chat transcript and to publish the transcript together with refined responses online where they can be accessed by attendees for reference. To help interested attendees, we offer consulting hours. Consulting hours are hosted as online telephone conferences where attendees can call in and ask further questions regarding lecture content or homework.

@@

A recent online course\cite{KOKKOS_LECTURE} has been publicly announced through social media and counted approximately 200 attendees. The majority of attendees are familiar with at least one parallel programming model and have a background in scientific computing. The listing in Figure~\ref{fig:lectureOutline} shows the structure of our online lectures.

Questions asked by attendees during the recent online course showed that they did understand the relevance as well as key concepts. Questions were mostly related to details in behavior, differences to other programming models, and best practices. We would like to direct the interested reader to our online website where the Q\&A sessions are listed in detail.
